Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-09 csf-2017/paper.tex

We define the syntax of our capability machine in Figure~\ref{fig:syntax}. We
assume an infinite set of addresses $\Addrs$ and define machine words as either
integers or capabilities of the form $((\perm,\gl),\start,\addrend,\addr)$. Such
a capability represents the authority to execute permissions $\perm$ on the
memory array $[\start,\addrend]$, together with a current address $\addr$ and a
tag $\gl$ which indicates whether the capability is global or local. The
available permissions and their ordering are depicted in
Figure~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
read/write/execute ($\rwx$) permissions. Additionally, there are three special
read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter
($\entry$) permissions which we will explain below.

@@

\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
\clearpage
}
\todo[inline]{Add section introduction}

@@

The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
differ only slightly. Executable capabilities can be updated to point
anywhere in memory, so they must be capability safe as a program
counter no matter where they point in memory. In contrast, enter
\lars{``capability safe as a program counter'' is a bit hard to
  understand - why not say that this is a capability one can jump to ?}
\lau{ When jumping to a capability, it becomes the program counter. A program counter is capability safe when it is the $\stder$ relation. When I say ``capability safe as a program counter'' I take it to mean being in the $\stder$-relation. }
capabilities are opaque and must be used from the address they point
to. They change permission when jumped to, so we require them to be
capability safe as a program counter when the permission is changed to
read-execute.
In both conditions, the capability must be capability safe as a
program counter in certain future worlds. If the capability is global,
then it should be capability safe as a program counter in any private
future world. This means that it can only depend on permanent regions
because temporary regions may be revoked in a private future
world. On the other hand, local capabilities must only be capability
safe in public future worlds because they may depend on temporary
regions, so we cannot expect them to function correctly as program
counters if temporary regions are revoked.

>>>>

We define the syntax of our capability machine in \figurename~\ref{fig:syntax}. We
assume an infinite set of addresses $\Addrs$ and define machine words as either
integers or capabilities of the form $((\perm,\gl),\start,\addrend,\addr)$. Such
a capability represents the authority to execute permissions $\perm$ on the
memory array $[\start,\addrend]$, together with a current address $\addr$ and a
tag $\gl$ which indicates whether the capability is global or local. The
available permissions and their ordering are depicted in
\figurename~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
read/write/execute ($\rwx$) permissions. Additionally, there are three special
read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter
($\entry$) permissions which we will explain below.

@@

\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
\clearpage
}
%subsection introduction
The logical relation defines semantically what it means to be
capability safe. In large parts, capability safety is defined by
when a capability safely can be given to an untrusted piece of
code. In the following paragraphs, we will go through the definitions
of the logical relation and provide an intuition why they define
capability safety.
The definition of the logical relation is found in
\figurename~\ref{fig:logrel}, \figurename~\ref{fig:standard-regions},
and \figurename~\ref{fig:perm-conds}. Some definitions are omitted and
explained when needed. The precise definitions can be found in the
technical report.\todo[inline]{add reference to technical report}

@@

The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
differ only slightly. Executable capabilities can be updated to point
anywhere in memory, so they must be capability safe as a program
counter no matter where they point in memory. In contrast, enter
\lars{``capability safe as a program counter'' is a bit hard to
  understand - why not say that this is a capability one can jump to ?}
\lau{ When jumping to a capability, it becomes the program counter. A program counter is capability safe when it is the $\stder$ relation. When I say ``capability safe as a program counter'' I take it to mean being in the $\stder$-relation. }
capabilities are opaque and must be used from the address they point
to. They change permission when jumped to, so we require them to be
capability safe as a program counter when the permission is changed to
read/execute.
In both conditions, it must be safe to jump to the capability. This
means that the capability must be capability safe as a
program counter in certain future worlds. If the capability is global,
then it should be capability safe as a program counter in any private
future world. This means that it can only depend on permanent regions
because temporary regions may be revoked in a private future
world. On the other hand, local capabilities must only be capability
safe in public future worlds because they may depend on temporary
regions, so we cannot expect them to function correctly as program
counters if temporary regions are revoked.
